---
title: "NYPD_Andree"
author: "Andree"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Introduction

```{r echo = FALSE}
library(tidyverse)
setwd("C:/Users/62812/OneDrive/Desktop/Andree/Colorado/DataScienceasaField/nypdProject")
nypdData <- read_csv("https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv")

nypdData$STATISTICAL_MURDER_FLAG_TRUE <- ifelse(nypdData$STATISTICAL_MURDER_FLAG == TRUE, 1, 0)
murder_p = round(100 *sum(nypdData$STATISTICAL_MURDER_FLAG_TRUE) / nrow(nypdData),2)
murder_class <- c("murder", "not murder")
murder_proportion <- c(murder_p, 100-murder_p)

pie(murder_proportion, labels = murder_proportion, 
    main = "Murder Proportion out of Shooting Occurrences %", col = rainbow(length(murder_proportion)))
legend("topright", c("murder", "not murder"), 
       cex = 0.5, fill = rainbow(length(murder_proportion)))

```




This document tries to predict if a shooting would result in death - hence counted as a murder - based on information such as: occur time of the day, boro, jurisdiction code, perpetrator's age, perpetrator's race, perpetrator's sex, victim's age, victim's race, victim's sex, victim's age. 

## Source

Raw data for this document is available at <"https://data.cityofnewyork.us/api/views/833y-fsy8/rows.csv"> for download.  
As described on NYC Open Data <"https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data">, the raw data contains every shooting occurences that happen between 2006 and 2023. 

"This data is manually extracted every quarter and reviewed by the Office of Management Analysis and Planning before being posted on the NYPD website. Each record represents a shooting incident in NYC and includes information about the event, the location and time of occurrence. In addition, information related to suspect and victim demographics is also included." - NYC Open Data


## Variables

```{r echo=FALSE}
nypdData$Boro_Manhattan <- ifelse(nypdData$BORO == "MANHATTAN", 1, 0)
nypdData$Boro_Bronx <- ifelse(nypdData$BORO == "BRONX", 1, 0)
nypdData$Boro_Queens <- ifelse(nypdData$BORO == "QUEENS", 1, 0)
nypdData$Boro_Brooklyn <- ifelse(nypdData$BORO == "BROOKLYN", 1, 0)
nypdData$Boro_StatenIsland <- ifelse(nypdData$BORO == "STATEN ISLAND", 1, 0)

boro_class <- c("Manhattan", "Bronx", "Queens", "Brooklyn", "Staten Island")

boro_occurrences_proportion <- c( round(sum(nypdData$Boro_Manhattan) * 100 / nrow(nypdData), 2), 
                                  round(sum(nypdData$Boro_Bronx) * 100/ nrow(nypdData), 2), 
                                  round(sum(nypdData$Boro_Queens) *100 / nrow(nypdData), 2), 
                                  round(sum(nypdData$Boro_Brooklyn) *100/ nrow(nypdData), 2), 
                                  round(sum(nypdData$Boro_StatenIsland)*100/ nrow(nypdData), 2))

pie(boro_occurrences_proportion, labels = boro_occurrences_proportion, 
    main = "Boro Occurrences %", col = rainbow(length(boro_occurrences_proportion)))
legend("topright", c("Manhattan", "Bronx", "Queens", "Brooklyn", "Staten Island"), 
       cex = 0.5, fill = rainbow(length(boro_occurrences_proportion)))
```

```{r echo=FALSE}
nypdData$OCCUR_TIME_5to10 <- ifelse(hour(nypdData$OCCUR_TIME) >= 5 & hour(nypdData$OCCUR_TIME) <= 10, 1, 0)
nypdData$OCCUR_TIME_10to15 <- ifelse(hour(nypdData$OCCUR_TIME) >= 10 & hour(nypdData$OCCUR_TIME) <= 15, 1, 0)
nypdData$OCCUR_TIME_15to20 <- ifelse(hour(nypdData$OCCUR_TIME) >= 15 & hour(nypdData$OCCUR_TIME) <= 20, 1, 0)
nypdData$OCCUR_TIME_20to24 <- ifelse(hour(nypdData$OCCUR_TIME) >= 20 & hour(nypdData$OCCUR_TIME) <= 24, 1, 0)
nypdData$OCCUR_TIME_00to5 <- ifelse(hour(nypdData$OCCUR_TIME) >= 00 & hour(nypdData$OCCUR_TIME) <= 5, 1, 0)


OccurTime_Window <- c("5 to 10", "10 to 15", "15 to 20", "20 to 24", "00 to 5")

OccurTime_proportion <- c( round(sum(nypdData$OCCUR_TIME_5to10) * 100 / nrow(nypdData), 2), 
                              round(sum(nypdData$OCCUR_TIME_10to15) * 100/ nrow(nypdData), 2),
                              round(sum(nypdData$OCCUR_TIME_15to20) * 100/ nrow(nypdData), 2),
                              round(sum(nypdData$OCCUR_TIME_20to24) * 100/ nrow(nypdData), 2),
                              round(sum(nypdData$OCCUR_TIME_00to5) * 100/ nrow(nypdData), 2))

pie(OccurTime_proportion, labels = OccurTime_proportion, 
    main = "Occur Time %", col = rainbow(length(OccurTime_proportion)))
legend("topright", c("5 to 10", "10 to 15", "15 to 20", "20 to 24", "00 to 5"), 
       cex = 0.5, fill = rainbow(length(OccurTime_proportion)))

```

```{r echo=FALSE}
nypdData$JURISDICTION_CODE_Patrol <- ifelse(is.na(nypdData$JURISDICTION_CODE) == TRUE, 0, ifelse(nypdData$JURISDICTION_CODE == "0", 1, 0))
nypdData$JURISDICTION_CODE_Transit <- ifelse(is.na(nypdData$JURISDICTION_CODE) == TRUE, 0, ifelse(nypdData$JURISDICTION_CODE == "1", 1, 0))
nypdData$JURISDICTION_CODE_Housing <- ifelse(is.na(nypdData$JURISDICTION_CODE) == TRUE, 0, ifelse(nypdData$JURISDICTION_CODE == "2", 1, 0))
nypdData$JURISDICTION_CODE_NA <- ifelse(is.na(nypdData$JURISDICTION_CODE) == TRUE, 1, 0)


JurisdictionGrp <- c("Patrol", "Transit", "Housing", "NA")

Jurisdiction_proportion <- c( round(sum(nypdData$JURISDICTION_CODE_Patrol) * 100 / nrow(nypdData), 2), 
                              round(sum(nypdData$JURISDICTION_CODE_Transit) * 100/ nrow(nypdData), 2),
                              round(sum(nypdData$JURISDICTION_CODE_Housing) * 100/ nrow(nypdData), 2),
                              round(sum(nypdData$JURISDICTION_CODE_NA) * 100/ nrow(nypdData), 2))


pie(Jurisdiction_proportion, labels = Jurisdiction_proportion, 
    main = "Jurisdiction %", col = rainbow(length(Jurisdiction_proportion)))
legend("topright", c("Patrol", "Transit", "Housing", "NA"), 
       cex = 0.5, fill = rainbow(length(Jurisdiction_proportion)))
       
```

```{r echo=FALSE}
nypdData$VicAgeGrp_2544 <- ifelse(nypdData$VIC_AGE_GROUP == "25-44", 1, 0)
nypdData$VicAgeGrp_1824 <- ifelse(nypdData$VIC_AGE_GROUP == "18-24", 1, 0)
nypdData$VicAgeGrp_bel18 <- ifelse(nypdData$VIC_AGE_GROUP == "<18", 1, 0)
nypdData$VicAgeGrp_ab65 <- ifelse(nypdData$VIC_AGE_GROUP == "65+", 1, 0)
nypdData$VicAgeGrp_4564 <- ifelse(nypdData$VIC_AGE_GROUP == "45-64", 1, 0)
nypdData$VicAgeGrp_NA <- ifelse(nypdData$VIC_AGE_GROUP == "UNKNOWN" | nypdData$VIC_AGE_GROUP == "1022", 1, 0)

VicAgeGrp <- c("< 18", "18 to 24", "25 to 44", "45 to 64", " > 65", "NA")
VicAgeGrp_proportion <- c( round(sum(nypdData$VicAgeGrp_bel18) * 100 / nrow(nypdData), 2), 
                           round(sum(nypdData$VicAgeGrp_1824) * 100/ nrow(nypdData), 2), 
                           round(sum(nypdData$VicAgeGrp_2544) *100 / nrow(nypdData), 2), 
                           round(sum(nypdData$VicAgeGrp_4564) *100/ nrow(nypdData), 2), 
                           round(sum(nypdData$VicAgeGrp_ab65)*100/ nrow(nypdData), 2),
                           round(sum(nypdData$VicAgeGrp_NA)*100/ nrow(nypdData), 2))

pie(VicAgeGrp_proportion, labels = VicAgeGrp_proportion, 
    main = "Victim Age Group %", col = rainbow(length(VicAgeGrp_proportion)))
legend("topright", c("<18", "18 to 24", "25 to 44", "45 to 64", "> 65", "NA"), 
       cex = 0.5, fill = rainbow(length(VicAgeGrp_proportion)))

```

```{r echo=FALSE}
nypdData$VicSex_Male <- ifelse(nypdData$VIC_SEX == "M", 1, 0)
nypdData$VicSex_NA <- ifelse(nypdData$VIC_SEX == "U", 1, 0)
nypdData$VicSex_Female <- ifelse(nypdData$VIC_SEX == "F", 1, 0)

VicSexGrp <- c("male", "female", "NA")

VicSexGrp_proportion <- c( round(sum(nypdData$VicSex_Male) * 100 / nrow(nypdData), 2), 
                           round(sum(nypdData$VicSex_Female) * 100/ nrow(nypdData), 2), 
                           round(sum(nypdData$VicSex_NA) *100 / nrow(nypdData), 2))

pie(VicSexGrp_proportion, labels = VicSexGrp_proportion, 
    main = "Victim Sex Group %", col = rainbow(length(VicSexGrp_proportion)))
legend("topright", c("male", "female", "NA"), 
       cex = 0.5, fill = rainbow(length(VicSexGrp_proportion)))
```


```{r echo=FALSE}
nypdData$VicRace_black <- ifelse(nypdData$VIC_RACE == "BLACK", 1, 0)
nypdData$VicRace_white <- ifelse(nypdData$VIC_RACE == "WHITE", 1, 0)
nypdData$VicRace_whiteHispanic <- ifelse(nypdData$VIC_RACE == "WHITE HISPANIC", 1, 0)
nypdData$VicRace_blackHispanic <- ifelse(nypdData$VIC_RACE == "BLACK HISPANIC", 1, 0)
nypdData$VicRace_asian <- ifelse(nypdData$VIC_RACE == "ASIAN / PACIFIC ISLANDER", 1, 0)
nypdData$VicRace_AmericanIndian <- ifelse(nypdData$VIC_RACE == "AMERICAN INDIAN/ALASKAN NATIVE", 1, 0)

VicRaceGrp <- c("Black", "White", "White Hispanic", "Black Hispanic", "Asian", "American Indian")

VicRaceGrp_proportion <- c( round(sum(nypdData$VicRace_black) * 100 / nrow(nypdData), 2), 
                           round(sum(nypdData$VicRace_white) * 100/ nrow(nypdData), 2), 
                           round(sum(nypdData$VicRace_whiteHispanic) *100 / nrow(nypdData), 2),
                           round(sum(nypdData$VicRace_blackHispanic) *100 / nrow(nypdData), 2),
                           round(sum(nypdData$VicRace_asian) *100 / nrow(nypdData), 2),
                           round(sum(nypdData$VicRace_AmericanIndian) *100 / nrow(nypdData), 2))

pie(VicRaceGrp_proportion, labels = VicRaceGrp_proportion, 
    main = "Victim Race Group %", col = rainbow(length(VicRaceGrp_proportion)))
legend("topright", c("Black", "White", "White Hispanic", "Black Hispanic", "Asian", "American Indian"), 
       cex = 0.5, fill = rainbow(length(VicRaceGrp_proportion)))

```


```{r echo=FALSE}
nypdData$PERP_AGE_GROUP_2544 <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 0, ifelse(nypdData$PERP_AGE_GROUP == "25-44", 1, 0))
nypdData$PERP_AGE_GROUP_1824 <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 0, ifelse(nypdData$PERP_AGE_GROUP == "18-24", 1, 0))
nypdData$PERP_AGE_GROUP_ab65 <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 0, ifelse(nypdData$PERP_AGE_GROUP == "65+", 1, 0))
nypdData$PERP_AGE_GROUP_NA <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 1, 0)
nypdData$PERP_AGE_GROUP_bel18 <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 0, ifelse(nypdData$PERP_AGE_GROUP == "<18", 1, 0))
nypdData$PERP_AGE_GROUP_4564 <- ifelse(nypdData$PERP_AGE_GROUP == "UNKNOWN" | nypdData$PERP_AGE_GROUP == "1022" | nypdData$PERP_AGE_GROUP == "940" | nypdData$PERP_AGE_GROUP == "224" | nypdData$PERP_AGE_GROUP == "1028" | is.na(nypdData$PERP_AGE_GROUP) == TRUE | nypdData$PERP_AGE_GROUP == "(null)", 0, ifelse(nypdData$PERP_AGE_GROUP == "45_64", 1, 0))

PERP_AGE_GROUP <- c("< 18", "18 to 24", "25 to 44", "45 to 64", " > 65", "NA")
PERP_AGE_GROUP_proportion <- c( round(sum(nypdData$PERP_AGE_GROUP_bel18) * 100 / nrow(nypdData), 2), 
                                round(sum(nypdData$PERP_AGE_GROUP_1824) * 100/ nrow(nypdData), 2), 
                                round(sum(nypdData$PERP_AGE_GROUP_2544) *100 / nrow(nypdData), 2), 
                                round(sum(nypdData$PERP_AGE_GROUP_4564) *100/ nrow(nypdData), 2), 
                                round(sum(nypdData$PERP_AGE_GROUP_ab65)*100/ nrow(nypdData), 2),
                                round(sum(nypdData$PERP_AGE_GROUP_NA)*100/ nrow(nypdData), 2))



pie(PERP_AGE_GROUP_proportion, labels = PERP_AGE_GROUP_proportion, 
    main = "Perpertrator Age Group %", col = rainbow(length(PERP_AGE_GROUP_proportion)))
legend("topright", c("<18", "18 to 24", "25 to 44", "45 to 64", "> 65", "NA"), 
       cex = 0.5, fill = rainbow(length(PERP_AGE_GROUP_proportion)))

```


```{r echo=FALSE}
nypdData$PERP_SEX_m <- ifelse(nypdData$PERP_SEX == "U" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_SEX == "M", 1, 0))
nypdData$PERP_SEX_NA <- ifelse(nypdData$PERP_SEX == "U" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 1, 0)
nypdData$PERP_SEX_F <- ifelse(nypdData$PERP_SEX == "U" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_SEX == "F", 1, 0))

PERP_SEX <- c("male", "female", "NA")

PERP_SEX_proportion <- c( round(sum(nypdData$PERP_SEX_m) * 100 / nrow(nypdData), 2), 
                          round(sum(nypdData$PERP_SEX_F) * 100/ nrow(nypdData), 2), 
                          round(sum(nypdData$PERP_SEX_NA) *100 / nrow(nypdData), 2))

pie(PERP_SEX_proportion, labels = PERP_SEX_proportion, 
    main = "Perpetrator Sex %", col = rainbow(length(PERP_SEX_proportion)))
legend("topright", c("male", "female", "NA"), 
       cex = 0.5, fill = rainbow(length(PERP_SEX_proportion)))

```

```{r echo=FALSE}
nypdData$PerpRace_black <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "BLACK", 1, 0)) 
nypdData$PerpRace_white <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "WHITE", 1, 0)) 
nypdData$PerpRace_whiteHispanic <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "WHITE HISPANIC", 1, 0)) 
nypdData$PerpRace_blackHispanic <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "BLACK HISPANIC", 1, 0)) 
nypdData$PerpRace_asian <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "ASIAN / PACIFIC ISLANDER", 1, 0)) 
nypdData$PerpRace_AmericanIndian <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_SEX == "(null)" | is.na(nypdData$PERP_SEX) == TRUE, 0, ifelse(nypdData$PERP_RACE == "AMERICAN INDIAN/ALASKAN NATIVE", 1, 0)) 
nypdData$PerpRace_NA <- ifelse(nypdData$PERP_RACE == "UNKNOWN" | nypdData$PERP_RACE == "(null)" | is.na(nypdData$PERP_RACE) == TRUE, 1, 0)



PerpRaceGrp <- c("Black", "White", "White Hispanic", "Black Hispanic", "Asian", "American Indian")

PerpRaceGrp_proportion <- c( round(sum(nypdData$PerpRace_black) * 100 / nrow(nypdData), 2), 
                             round(sum(nypdData$PerpRace_white) * 100/ nrow(nypdData), 2), 
                             round(sum(nypdData$PerpRace_whiteHispanic) *100 / nrow(nypdData), 2),
                             round(sum(nypdData$PerpRace_blackHispanic) *100 / nrow(nypdData), 2),
                             round(sum(nypdData$PerpRace_asian) *100 / nrow(nypdData), 2),
                             round(sum(nypdData$PerpRace_AmericanIndian) *100 / nrow(nypdData), 2))

sum(nypdData$PerpRace_black)

pie(PerpRaceGrp_proportion, labels = PerpRaceGrp_proportion, 
    main = "Perpetrator's Race Group %", col = rainbow(length(PerpRaceGrp_proportion)))
legend("topright", c("Black", "White", "White Hispanic", "Black Hispanic", "Asian", "American Indian"), 
       cex = 0.5, fill = rainbow(length(PerpRaceGrp_proportion)))

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



## Model 1

Assuming independence, in theory, we could calculate the joint probability of death in a shooting given the variables above.

### Analysis

Assuming independence between the above predictors, an 11-am shooting in Brooklyn of a 19yo male White Hispanic Victim by a 22yo male White Hispanic perpetrator in a patrol jurisdiction would result in death - hence categorized as murder - by the following probability formula: 

```{r echo=TRUE}
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & hour(nypdData$OCCUR_TIME) >= 10 & hour(nypdData$OCCUR_TIME) <= 15 )) / sum(nypdData$OCCUR_TIME_10to15) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$BORO == "BROOKLYN"))/ sum(nypdData$Boro_Brooklyn)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_SEX == "M" )) /sum(nypdData$VicSex_Male) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_RACE == "WHITE HISPANIC"))/ sum(nypdData$VicRace_whiteHispanic)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_AGE_GROUP == "18-24")) / sum(nypdData$VicAgeGrp_1824) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_AGE_GROUP == "18-24")) / sum(nypdData$PERP_AGE_GROUP_1824))*
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_SEX == "M")) / sum(nypdData$PERP_SEX_m))*
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_RACE == "WHITE HISPANIC")) / sum(nypdData$PerpRace_whiteHispanic)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$JURISDICTION_CODE == "0")) / sum(nypdData$JURISDICTION_CODE_Patrol)) 
```


### Comparing with Real Data

In reality, an 11-am shooting in Brooklyn of a 19yo male White Hispanic Victim by a 22yo male White Hispanic perpetrator in a patrol jurisdiction occurs 0.01% out of the shooting occurrences.

```{r echo=TRUE}

testcase <- nypdData %>%
              filter(
                hour(nypdData$OCCUR_TIME) >= 10 & hour(nypdData$OCCUR_TIME) <= 15 &
                nypdData$BORO == "BROOKLYN" &
                nypdData$VIC_SEX == "M" & 
                nypdData$VIC_AGE_GROUP == "18-24" &
                nypdData$VIC_RACE == "WHITE HISPANIC" &
                nypdData$PERP_AGE_GROUP == "18-24" &
                nypdData$PERP_SEX == "M" &
                nypdData$PERP_RACE == "WHITE HISPANIC" &
                nypdData$JURISDICTION_CODE == "0" )


nrow(testcase) *100 / nrow(nypdData)

```

The two numbers are very far away from another. 

Let's try another case of, say, a 10-pm shooting in Bronx of a 30yo male Black Victim by a 17yo male Black perpetrator in a housing jurisdiction.

```{r echo=TRUE}
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & hour(nypdData$OCCUR_TIME) >= 20 & hour(nypdData$OCCUR_TIME) <= 24 )) / sum(nypdData$OCCUR_TIME_20to24) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$BORO == "BRONX"))/ sum(nypdData$Boro_Bronx)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_SEX == "M" )) /sum(nypdData$VicSex_Male) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_RACE == "BLACK"))/ sum(nypdData$VicRace_black)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$VIC_AGE_GROUP == "25-44")) / sum(nypdData$VicAgeGrp_2544) ) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_AGE_GROUP == "<18")) / sum(nypdData$PERP_AGE_GROUP_bel18))*
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_SEX == "M")) / sum(nypdData$PERP_SEX_m))*
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$PERP_RACE == "BLACK")) / sum(nypdData$PerpRace_black)) *
  (nrow(nypdData %>% filter(nypdData$STATISTICAL_MURDER_FLAG == TRUE & nypdData$JURISDICTION_CODE == "2")) / sum(nypdData$JURISDICTION_CODE_Housing)) 
```

```{r echo=TRUE}

testcase2 <- nypdData %>%
              filter(
                hour(nypdData$OCCUR_TIME) >= 20 & hour(nypdData$OCCUR_TIME) <= 24 &
                nypdData$BORO == "BRONX" &
                nypdData$VIC_SEX == "M" & 
                nypdData$VIC_AGE_GROUP == "25-44" &
                nypdData$VIC_RACE == "BLACK" &
                nypdData$PERP_AGE_GROUP == "<18" &
                nypdData$PERP_SEX == "M" &
                nypdData$PERP_RACE == "BLACK" &
                nypdData$JURISDICTION_CODE == "2" )


nrow(testcase2) *100 / nrow(nypdData)

```

They too are far away from one another..

### Conclusion
I cannot confidently say that the above joint probability model is accurate. 


### Possible Bias
The model above does not remove nor omit NA values. 


## Model 2 - Logistic Regression

A quick googling would tell us that one of the best ways to model discrete outcomes is logistic regression.
Let's attempt to set one up using the NYPD data.



Also, while we are at it, let's filter out the NAs

```{r echo=TRUE}

nypdModel <- nypdData %>%
  filter( VicAgeGrp_NA == 0 & VicSex_NA == 0 & JURISDICTION_CODE_NA == 0  & PERP_AGE_GROUP_NA == 0 & PERP_SEX_NA == 0 & nypdData$PerpRace_NA == 0) %>%
  select(INCIDENT_KEY, STATISTICAL_MURDER_FLAG_TRUE, 
         Boro_Bronx, Boro_Manhattan, Boro_Queens, Boro_Brooklyn, Boro_StatenIsland,
         VicAgeGrp_2544, VicAgeGrp_1824, VicAgeGrp_bel18, VicAgeGrp_ab65, VicAgeGrp_4564,
         VicSex_Male, VicSex_Female,
         VicRace_black, VicRace_white, VicRace_whiteHispanic, VicRace_blackHispanic, VicRace_asian, VicRace_AmericanIndian,
         JURISDICTION_CODE_Transit, JURISDICTION_CODE_Patrol, JURISDICTION_CODE_Housing,
         OCCUR_TIME_5to10, OCCUR_TIME_10to15, OCCUR_TIME_15to20, OCCUR_TIME_20to24,OCCUR_TIME_00to5,
         PerpRace_black, PerpRace_white, PerpRace_whiteHispanic, PerpRace_blackHispanic, PerpRace_asian, PerpRace_AmericanIndian,
         PERP_AGE_GROUP_2544, PERP_AGE_GROUP_1824, PERP_AGE_GROUP_ab65, PERP_AGE_GROUP_bel18, PERP_AGE_GROUP_4564,
         PERP_SEX_m, PERP_SEX_F)

```


To avoid collinearity, we leave out one classification for each predictors; i.e., we would leave: Staten Island for Boro, 00 to 5 for occur time, etc.  
It is to be noted that the following codes are inspired by the following blog <"https://www.geeksforgeeks.org/filter-data-by-multiple-conditions-in-r-using-dplyr/">

```{r echo=TRUE}
mod2 <- glm(STATISTICAL_MURDER_FLAG_TRUE ~ Boro_Brooklyn + Boro_Queens + Boro_Bronx + Boro_Manhattan +
          VicAgeGrp_1824+ VicAgeGrp_2544 + VicAgeGrp_ab65 + VicAgeGrp_bel18 +
          VicSex_Male +
          VicRace_black + VicRace_white + VicRace_whiteHispanic + VicRace_blackHispanic + VicRace_asian +
          JURISDICTION_CODE_Transit + JURISDICTION_CODE_Patrol +
          OCCUR_TIME_5to10 + OCCUR_TIME_10to15 + OCCUR_TIME_15to20 + OCCUR_TIME_20to24 +
          PERP_AGE_GROUP_2544 + PERP_AGE_GROUP_1824 + PERP_AGE_GROUP_ab65 + PERP_AGE_GROUP_bel18 +
          PerpRace_black + PerpRace_white + PerpRace_whiteHispanic + PerpRace_blackHispanic + PerpRace_asian +
          PERP_SEX_m,
          family = binomial,
          data = nypdModel)

summary(mod2)

pred2 = predict(mod2)
summary(pred2)

```

I have to be honest: I do not yet know how to interpret these summaries. 
But I suspect, to have a gist on the model's performance, we could compare its predictions and the actual result.

First we could find out the proportion of murder among the shooting occurrences. 
```{r echo=TRUE}

sum(nypdModel$STATISTICAL_MURDER_FLAG_TRUE)/nrow(nypdModel)
```

When the NAs have been filtered out, occurrences were classified as murder ~23.93% of the time. 

Let's look at how often the model predicts murder:

```{r echo=TRUE}

sum(pred2)/nrow(nypdModel)

```

The result is a weird_looking negative number. 



As everything is already set up, let's try to also evaluate the performance of model1 with data that does not contain NAs. 

```{r echo = FALSE}
Boro_Brooklyn_Proportion = sum(nypdModel$Boro_Brooklyn) / nrow(nypdModel)
Boro_Queens_Proportion = sum(nypdModel$Boro_Queens) / nrow(nypdModel)
Boro_Bronx_Proportion = sum(nypdModel$Boro_Bronx) / nrow(nypdModel)
Boro_Manhattan_Proportion = sum(nypdModel$Boro_Manhattan) / nrow(nypdModel)
Boro_StatenIsland_Proportion = 1 - (Boro_Manhattan_Proportion + Boro_Bronx_Proportion + Boro_Queens_Proportion + Boro_Brooklyn_Proportion)

VicAgeGrp_1824_Proportion = sum(nypdModel$VicAgeGrp_1824) / nrow(nypdModel)
VicAgeGrp_2544_Proportion = sum(nypdModel$VicAgeGrp_2544) / nrow(nypdModel)
VicAgeGrp_ab65_Proportion = sum(nypdModel$VicAgeGrp_ab65) / nrow(nypdModel)
VicAgeGrp_bel18_Proportion = sum(nypdModel$VicAgeGrp_bel18) / nrow(nypdModel)
VicAgeGrp_4564_Proportion = 1 - (VicAgeGrp_1824_Proportion + VicAgeGrp_bel18_Proportion + VicAgeGrp_2544_Proportion + VicAgeGrp_ab65_Proportion)

VicMale_Proportion = sum(nypdModel$VicSex_Male) / nrow(nypdModel)
VicFemale_Proportion = 1- VicMale_Proportion

VicRace_black_Proportion = sum(nypdModel$VicRace_black) / nrow(nypdModel)
VicRace_white_Proportion = sum(nypdModel$VicRace_white) / nrow(nypdModel)
VicRace_whiteHispanic_Proportion = sum(nypdModel$VicRace_whiteHispanic) / nrow(nypdModel)
VicRace_blackHispanic_Proportion = sum(nypdModel$VicRace_blackHispanic) / nrow(nypdModel)
VicRace_asian_Proportion = sum(nypdModel$VicRace_asian) / nrow(nypdModel)
VicRace_americanIndian_Proportion = 1 - (VicRace_black_Proportion + VicRace_white_Proportion + VicRace_whiteHispanic_Proportion + VicRace_blackHispanic_Proportion + VicRace_asian_Proportion)

JURISDICTION_CODE_Transit_P = sum(nypdModel$JURISDICTION_CODE_Transit) / nrow(nypdModel)
JURISDICTION_CODE_Patrol_P = sum(nypdModel$JURISDICTION_CODE_Patrol) / nrow(nypdModel)
JURISDICTION_CODE_House_P = 1 - JURISDICTION_CODE_Patrol_P - JURISDICTION_CODE_Transit_P

OCCUR_TIME_5to10_P = sum(nypdModel$OCCUR_TIME_5to10)/ nrow(nypdModel)
OCCUR_TIME_10to15_P = sum(nypdModel$OCCUR_TIME_10to15)/ nrow(nypdModel)
OCCUR_TIME_15to20_P = sum(nypdModel$OCCUR_TIME_15to20)/ nrow(nypdModel)
OCCUR_TIME_20to24_P = sum(nypdModel$OCCUR_TIME_20to24)/ nrow(nypdModel)
OCCUR_TIME_00to5_P = 1 - (OCCUR_TIME_20to24_P + OCCUR_TIME_15to20_P + OCCUR_TIME_10to15_P + OCCUR_TIME_5to10_P )

PERP_AGE_GROUP_1824_Proportion = sum(nypdModel$PERP_AGE_GROUP_1824) / nrow(nypdModel)
PERP_AGE_GROUP_2544_Proportion = sum(nypdModel$PERP_AGE_GROUP_2544) / nrow(nypdModel)
PERP_AGE_GROUP_ab65_Proportion = sum(nypdModel$PERP_AGE_GROUP_ab65) / nrow(nypdModel)
PERP_AGE_GROUP_bel18_Proportion = sum(nypdModel$PERP_AGE_GROUP_bel18) / nrow(nypdModel)
PERP_AGE_GROUP_4564_Proportion = 1 - (PERP_AGE_GROUP_1824_Proportion + PERP_AGE_GROUP_bel18_Proportion + PERP_AGE_GROUP_2544_Proportion + PERP_AGE_GROUP_ab65_Proportion)

PerpMale_Proportion = sum(nypdModel$PERP_SEX_m) / nrow(nypdModel)
PerpFemale_Proportion = 1- PerpMale_Proportion

PerpRace_black_Proportion = sum(nypdModel$PerpRace_black) / nrow(nypdModel)
PerpRace_white_Proportion = sum(nypdModel$PerpRace_white) / nrow(nypdModel)
PerpRace_whiteHispanic_Proportion = sum(nypdModel$PerpRace_whiteHispanic) / nrow(nypdModel)
PerpRace_blackHispanic_Proportion = sum(nypdModel$PerpRace_blackHispanic) / nrow(nypdModel)
PerpRace_asian_Proportion = sum(nypdModel$PerpRace_asian) / nrow(nypdModel)
PerpRace_americanIndian_Proportion = 1 - (PerpRace_black_Proportion + PerpRace_white_Proportion + PerpRace_whiteHispanic_Proportion + PerpRace_blackHispanic_Proportion + PerpRace_asian_Proportion)


```

```{r echo=TRUE}


nypdModel$mod1 <- ((nypdModel$Boro_Bronx * Boro_Bronx_Proportion) + 
                      (nypdModel$Boro_Manhattan * Boro_Manhattan_Proportion) + 
                      (nypdModel$Boro_Queens * Boro_Queens_Proportion) + 
                      (nypdModel$Boro_Brooklyn * Boro_Brooklyn_Proportion) + 
                      (nypdModel$Boro_StatenIsland * Boro_StatenIsland_Proportion)) *
  
                    ((nypdModel$VicAgeGrp_1824 * VicAgeGrp_1824_Proportion) + 
                      (nypdModel$VicAgeGrp_2544 * VicAgeGrp_2544_Proportion) + 
                      (nypdModel$VicAgeGrp_4564 * VicAgeGrp_4564_Proportion) + 
                      (nypdModel$VicAgeGrp_ab65 * VicAgeGrp_ab65_Proportion) + 
                      (nypdModel$VicAgeGrp_bel18 * VicAgeGrp_bel18_Proportion)) *
  
                    ((nypdModel$VicSex_Male * VicMale_Proportion) + 
                      (nypdModel$VicSex_Female * VicFemale_Proportion)) *
  
                    ((nypdModel$VicRace_AmericanIndian * VicRace_americanIndian_Proportion) + 
                       (nypdModel$VicRace_asian * VicRace_asian_Proportion) + 
                       (nypdModel$VicRace_blackHispanic * VicRace_blackHispanic_Proportion) + 
                       (nypdModel$VicRace_whiteHispanic * VicRace_whiteHispanic_Proportion) + 
                       (nypdModel$VicRace_white * VicRace_white_Proportion) + 
                       (nypdModel$VicRace_black * VicRace_black_Proportion)) *
  
                    ((nypdModel$JURISDICTION_CODE_Patrol * JURISDICTION_CODE_Patrol_P) + 
                       (nypdModel$JURISDICTION_CODE_Housing * JURISDICTION_CODE_House_P) + 
                       (nypdModel$JURISDICTION_CODE_Transit * JURISDICTION_CODE_Transit_P)) *
  
                    ((nypdModel$OCCUR_TIME_00to5 * OCCUR_TIME_00to5_P) + 
                       (nypdModel$OCCUR_TIME_5to10 * OCCUR_TIME_5to10_P) + 
                       (nypdModel$OCCUR_TIME_10to15 * OCCUR_TIME_10to15_P) + 
                       (nypdModel$OCCUR_TIME_15to20 * OCCUR_TIME_15to20_P) + 
                       (nypdModel$OCCUR_TIME_20to24 * OCCUR_TIME_20to24_P)) *
  
                    ((nypdModel$PERP_AGE_GROUP_1824 * PERP_AGE_GROUP_1824_Proportion) + 
                       (nypdModel$PERP_AGE_GROUP_2544 * PERP_AGE_GROUP_2544_Proportion) + 
                       (nypdModel$PERP_AGE_GROUP_4564 * PERP_AGE_GROUP_4564_Proportion) + 
                       (nypdModel$PERP_AGE_GROUP_ab65 * PERP_AGE_GROUP_ab65_Proportion) + 
                       (nypdModel$PERP_AGE_GROUP_bel18 * PERP_AGE_GROUP_bel18_Proportion)) *
  
                    ((nypdModel$PERP_SEX_m * PerpMale_Proportion) + 
                       (nypdModel$PERP_SEX_F * PerpFemale_Proportion)) *
  
                    ((nypdModel$PerpRace_AmericanIndian * PerpRace_americanIndian_Proportion) + 
                       (nypdModel$PerpRace_asian * PerpRace_asian_Proportion) + 
                       (nypdModel$PerpRace_blackHispanic * PerpRace_blackHispanic_Proportion) + 
                       (nypdModel$PerpRace_whiteHispanic * PerpRace_whiteHispanic_Proportion) + 
                       (nypdModel$PerpRace_white * PerpRace_white_Proportion) + 
                       (nypdModel$PerpRace_black * PerpRace_black_Proportion))
  

nypdModel$pred1 <- ifelse(nypdModel$mod1 > 0.5, 1, 0)

sum(nypdModel$pred1)/nrow(nypdModel)


```

The result is also a suspicious zero.

Or may be, we should be summing mod1, instead of summing pred1..

```{r echo=TRUE}
sum(nypdModel$mod1)/nrow(nypdModel)

```



That is less suspicious. But can I do that?
Where did I do wrong?
Let me know what you think.


### Conclusion
The author requires more in-depth understanding of data science to model a project this ambitious.


### Possible Bias
The model above does not include rows with NAs among the selected predictors.
This could mean that the model is selectively bias as the NAs could be huge underrepresentations. 




